---
title: "Final Project"
author: "Alyssa Mondelli"
date: "`r Sys.Date()`"
output: 
  html_document:
    df_print: paged
    code_download: yes
---

```{r setup, include=FALSE}
bobross <- read.csv(file = file.choose(), header = TRUE)
library(dplyr)
library(ggplot2)
library(tidyr)
library(forcats)
library(gt)
library(paletteer)
library(gtExtras)
```

## Introduction to Dataset

My dataset is on the paintings of Bob Ross and is available from the [Tidy Tuesday](https://github.com/rfordatascience/tidytuesday/tree/master/data/2023/2023-02-21) github site. This dataset was created by Jared Wilber and contains observations about 403 paintings created over 31 seasons of Bob Ross's show The Joy of Painting including painting titles, what episode they were created in, and what colors and combinations of colors were used in each painting.

Questions:

(@) What is the general distribution of number of colors used in a Bob Ross painting?

(@) Does the average number of colors used vary across seasons of the show?

(@) What color scheme did Bob use in any given episode?

(@) Does Bob have favorite colors?

## Results

#### **Question 1:** What is the general distribution of number of colors used in a Bob Ross painting?

```{r message=FALSE, echo=FALSE}
bobross %>%
  ggplot(aes(x=num_colors, fill="brown")) +
  guides(fill="none") +
  geom_histogram() +
  theme_bw() +
  xlab("Number of Colors Used") +
  ylab("Number of Paintings") +
  labs(title=" Overall Distribution of Number of Colors Used in a Bob Ross Painting", caption="Source: Tidy Tuesday Bob Ross dataset from 2/21/2023")
```

This histogram shows the distribution of number of colors used per painting across all 31 seasons. While Bob has used as few as 1 and as many as 15 colors, the typical painting contains between 10 to 13 colors.


#### **Question 2:** Does the average number of colors used vary across seasons of the show?


```{r echo=FALSE, message=FALSE}

bobross %>%
  select(season, episode, num_colors) %>%
  group_by(season) %>%
  summarize(mean_colors=mean(num_colors)) %>%
  arrange(desc(mean_colors)) %>%
  ggplot(aes(y=mean_colors, x=season)) +
  guides(color="none") +
  geom_line(color="deepskyblue4") +
  labs(title="Mean Colors Used Across Seasons 1-31", caption="Source: Tidy Tuesday Bob Ross dataset from 2/21/2023") +
  xlab("Season") +
  ylab("Mean Colors Used") +
  theme_classic()
```

There is a surprising amount of variation between seasons. Season 1 was an outlier with a mean of just over 7 colors used. In a typical season Bob uses a mean of about 10 to 11.5 colors per painting.


#### **Question 3:** What color scheme did Bob use in any given episode?
```{r echo=FALSE}
bobross %>%
  pivot_longer(cols=c(Black_Gesso:Alizarin_Crimson), names_to="colornames", values_to="colorused") %>%
  filter(colorused != FALSE) %>%
  mutate(colornames = recode(colornames, Alizarin_Crimson = "Alizarin Crimson", Black_Gesso = "Black Gesso", Bright_Red = "Bright Red", Cadmium_Yellow = "Cadmium Yellow", Phthalo_Green = "Phthalo Green", Prussian_Blue = "Prussian Blue", Sap_Green = "Sap Green", Titanium_White = "Titanium White", Van_Dyke_Brown = "Van Dyke Brown")) %>%
  filter(season==1)%>%
  select(season, episode, colornames, colorused) %>%
  ggplot(aes(x=episode, fill=colornames)) +
  geom_bar() +
  labs(title="Color Scheme Used by Episode, Season 1", caption="Source: Tidy Tuesday Bob Ross dataset from 2/21/2023") +
  xlab("Episode") +
  ylab("Number of Colors Used") +
  theme_light()+ 
    scale_fill_manual(values=c("#4e1500", "#000000", "#db0000", "gold1", "seagreen", "royalblue4", "yellowgreen", "whitesmoke", "burlywood4"), name="")
```

As an example, here are the color schemes used in season 1. Bob used a total of nine colors during that season but only used all of them together once, in episode 3. Episode 4 used the fewest colors, only Prussian Blue, Titanium White, and Van Dyke Brown.

```{r echo=FALSE}
bobross %>%
  pivot_longer(cols=c(Black_Gesso:Alizarin_Crimson), names_to="colornames", values_to="colorused") %>%
  filter(colorused != FALSE) %>%
  mutate(colornames = recode(colornames, Alizarin_Crimson = "Alizarin Crimson", Black_Gesso = "Black Gesso", Bright_Red = "Bright Red", Burnt_Umber = "Burnt Umber", Cadmium_Yellow = "Cadmium Yellow", Dark_Sienna = "Dark Sienna", Indian_Red = "Indian Red", Indian_Yellow = "Indian Yellow", Liquid_Black = "Liquid Black", Liquid_Clear = "Liquid Clear", Midnight_Black = "Midnight Black", Phthalo_Blue = "Phthalo Blue", Phthalo_Green = "Phthalo Green", Prussian_Blue = "Prussian Blue", Sap_Green = "Sap Green", Titanium_White = "Titanium White", Van_Dyke_Brown = "Van Dyke Brown", Yellow_Ochre = "Yellow Ochre")) %>%
  select(season, episode, colornames, colorused) %>%
  ggplot(aes(x=episode, fill=colornames)) +
  geom_bar() +
  facet_wrap(~season) +
  labs(title="Color Scheme Used by Episode, Seasons 1-31", caption="Source: Tidy Tuesday Bob Ross dataset from 2/21/2023") +
  xlab("Episode") +
  ylab("Number of Colors Used") + 
    scale_fill_manual(values=c("#4e1500", "#000000", "#db0000", "darkorange4", "gold1", "sienna4", "indianred4", "orange1", "gray1", "ghostwhite", "gray2","dodgerblue4", "seagreen", "royalblue4", "yellowgreen", "whitesmoke", "burlywood4", "darkgoldenrod1"),name="")
  
```

We can extend this visualization to show the color schemes used across all 31 seasons.


#### **Question 4:** Does Bob have favorite colors?
```{r echo=FALSE, message=FALSE, warning=FALSE}
bobross %>%
  pivot_longer(cols=c(Black_Gesso:Alizarin_Crimson), names_to="colornames", values_to="colorused") %>%
  select(season, episode, colornames, colorused) %>%
  group_by(colornames) %>%
  mutate(colornames = recode(colornames, Alizarin_Crimson = "Alizarin Crimson", Black_Gesso = "Black Gesso", Bright_Red = "Bright Red", Burnt_Umber = "Burnt Umber", Cadmium_Yellow = "Cadmium Yellow", Dark_Sienna = "Dark Sienna", Indian_Red = "Indian Red", Indian_Yellow = "Indian Yellow", Liquid_Black = "Liquid Black", Liquid_Clear = "Liquid Clear", Midnight_Black = "Midnight Black", Phthalo_Blue = "Phthalo Blue", Phthalo_Green = "Phthalo Green", Prussian_Blue = "Prussian Blue", Sap_Green = "Sap Green", Titanium_White = "Titanium White", Van_Dyke_Brown = "Van Dyke Brown", Yellow_Ochre = "Yellow Ochre")) %>%
  summarize(total_colorused = sum(colorused)) %>%
  arrange(desc(total_colorused)) %>%
  gt() %>%
  cols_label(colornames = "Color Name", total_colorused = "Times Used") %>%
  tab_header(title = "Paint Colors Used by Bob Ross",
             subtitle = "n=403 paintings") %>%
  tab_source_note(source_note = "Source: Tidy Tuesday Bob Ross dataset from 2/21/2023") %>%
  gt_color_rows(columns=c("total_colorused"), 
                palette="futurevisions::cancri", domain = NULL,
                pal_type=c("continuous"))

```

After Titanium White, which is used in nearly every painting, Bob's next five most used colors are Alizarin Crimson, Van Dyke Brown, Cadmium Yellow, Yellow Ochre, and Phthalo Blue. His least used color, Indian Red, has only appeared in a single painting.


## Conclusions and Future Work

A typical Bob Ross painting contains between 10 to 13 colors, with a minimum of 1 and a maximum of 15. The distribution of number of colors used varies widely across and between seasons, but the mean number of colors used per season is typically between 10 and 11.5 colors per painting. Bob's five most frequently used colors (after white) are Alizarin Crimson, Van Dyke Brown, Cadmium Yellow, Yellow Ochre, and Phthalo Blue.

Additional questions: Do colors "travel together", aka are there certain combinations of color that Bob uses more often than others? Does the combination of colors used vary by time of year? (This could be answered if we had additional data about when each episode was recorded.) 

## References
<https://github.com/rfordatascience/tidytuesday/tree/master/data/2023/2023-02-21>